---
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


#Statistical Inference Project 1
##Exploring Asymptotics of the Exponential Distribution
###Chris Jones

In this document I will investigate the exponential distribution in R and compare it with the Central Limit Theorem.

The Central Limit Theorem states that given a distribution with mean $mu$ and variance $\sigma^{2}$, the sampling distribution of the mean approaches a normal distribution with mean $mu$ and variance $\frac {\sigma^{2}}{N}$. Counter-intuitively, the sampling distribution of the mean approaches a normal distribution regardless of the shape of the original distribution.


```{r}
library(ggplot2) #load the required packages
```

Let's set the parameters:

```{r}
lambda <- 0.2 #the rate of the exponential distribution
mu <- 1/lambda #the mean of the exponential distribution
n <- 40 #numbers of exponential numbers to be drawn
nsim <- 1000 #numbers of simulations to be performed
```

## Sample and Theoretical Mean

The theoretical mean of an exponential distribution with rate $\lambda$ is:

$\mu= \frac{1}{\lambda}$ 

```{r}
mu <- 1/lambda
mu
```

Let $\bar X$ be the sample mean of 1000 sets of 40 random exponential distributions with $\lambda = 0.2$

```{r}
set.seed(144) #set the seed for reproducibility

expDist <- matrix(data= NA, nrow = nsim, ncol = n) #empty matrix with 1000 rows and 40 columns
for(i in 1:nsim) {#for each row of the matrix,
	expDist[i,] <- rexp(n = n, rate = lambda) #draw 40 random numbers from the exponential distribution
}
means <- rowMeans(expDist) 
xbar <- mean(means)
xbar
```

$\bar X$ and $\mu$ are almost identical, as the Central Limit Theorem predicts.

## Sample and Theoretical Variance

The expected variance  of our sample of 1000 means is $\frac{\frac{1}{\lambda^{2}}}{40} = 0.625$. 

```{r}
var(means)
```

The theoretical and the sample variance are very similar. This matches the expectations of the Central Limit Theorem.

## Distribution

The Central Limit Theorem states that the means of our 1000 simulations should approximate a distribution:  $N(\frac{1}{0.2}, \frac{\frac{1}{0.2}}{\sqrt{40}})$. The histogram

```{r echo = FALSE}
df <- as.data.frame(means - mu)
ggplot(data = df) + 
	aes(x = means) +
	geom_histogram(binwidth = 0.4, color = "black", fill = "white", aes(y=..density..)) +
	stat_function(aes(x = c(1.5, 8.5)), fun = dnorm, arg = list(mean = 5 , sd = sqrt(0.625)), colour = "red", size = 1) +
	xlab("Sample mean") +
	ylab("Density") +
	ggtitle("Sample distribution vs. theoretical distribution")
```

Again, our sample and theoretical distributions are very close to one another.
